Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-10-31 forecast/paper/paper.tex
Copyedits to discussion

However, rigorously vetting the details for more sophisticated models, as well
as the generalizability of our findings to different geographic resolutions,
both remain important directions for future study.  For example, in the
Supplementary Materials we show that for forecasting at the state level, the
benefits of adding indicators to the AR model are generally less clear
(compared to those observed at the HRR level).  A plausible explanation: at the
state level, where the signal-to-noise ratio (SNR) is higher, AR performs better
overall and provides a higher standard (when asking whether it can be improved 
upon using indicators).  At the HRR level, where the SNR is lower, including the
indicators as additional linear features in the AR model probably delivers a
kind of variance reduction (just like averaging independent terms would) which 
contributes to improved accuracy.  But at a higher SNR, this variance
reduction becomes less important, and we may be compelled to use more
sophisticated modeling techniques to extract comparable value from the
indicators.

>>>>

However, rigorously vetting the details for more sophisticated models, as well
as the generalizability of our findings to different geographic resolutions,
both remain important directions for future study.  For example, in the
Supplementary Materials we show that for forecasting at the state level, the
benefits of including indicators in the AR model are generally less clear
(compared to those observed at the HRR level).  A plausible explanation: at the
state level, where the signal-to-noise ratio (SNR) is higher, AR performs better
overall and represents a higher standard (when asking whether it can be improved
upon using the indicators).  At the HRR level, where the SNR is lower, including
the indicators as additional linear features in the AR model probably delivers a
kind of variance reduction (just like averaging independent terms would) which
contributes to improved accuracy.  But as the SNR increases, this variance
reduction becomes less important, and perhaps we must use more sophisticated
modeling techniques to extract comparable value from the indicators.

### 2021-10-31 forecast/paper/supplement-text.tex
Copyedits to discussion

As explained in Section 2.B of the main paper, a (somewhat cynical) hypothesis
for why we see benefits in forecasting and hotspot prediction is that the
indicators are not actually providing useful information but they are instead
acting as a sort of ``implicit regularization,'' leading to shrinkage on the
autoregressive coefficients and therefore to less volatile predictions.  To
investigate this hypothesis, we consider fitting ``noise features'' that in
truth have zero relationship to the response.  Recall (from the main paper) that
at each forecast date, we train a model on 6,426 location-time pairs.  Indicator
models are based on six features, corresponding to the three autoregressive
terms and the three lagged indicator values.  To form noise indicator features,
we replace their values with those from a randomly chosen time-space pair (while
keeping the autoregressive features fixed).  In particular, at each location
$\ell$ and time $t$, for the forecasting task we replace the triplet
\smash{$(X_{\ell,t}, X_{\ell,t-7}, X_{\ell,t-14})$} in Eq.\ (3) of the main
paper with the triplet \smash{$(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
  X_{\ell^*,t^*-14})$}, where $(\ell^*,t^*)$ is a location-time pair sampled with
replacement from the 6,426 total location-time pairs.  Likewise in the hotspot 
prediction task, we replace the triplet \smash{$(X_{\ell,t}^\Delta,
  X_{\ell,t-7}^\Delta, X_{\ell,t-14}^\Delta)$} in Eq.\ (5) of the main paper with
\smash{$(X_{\ell^*,t^*}^\Delta, X_{\ell^*,t^*-7}^\Delta,
  X_{\ell^*,t^*-14}^\Delta)$}.  Figures
\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No method
exhibits a noticeable performance gain over the \ar~method, leading us to
dismiss the implicit regularization hypothesis.

@@

In this section, we provide extra details about the up/down/flat analysis
described in the main text. Figure \ref{fig:upswing-summary} shows the overall
results, examining the average difference $\WIS(AR) - \WIS(F)$ in period. Figure
\ref{fig:hotspots-upswing-downswing} shows the same information for the hotspot
task. On average, during downswings and flat periods, the indicator models have
lower classification error and higher log likelihood than the AR model. For
hotspots, both Google-AA and CTIS-CLIIC perform better than the AR model during
upswings, in contrast to the forecasting task, where only Google-AA
improves. For a related analysis, Figure \ref{fig:cor-wis-ratio} shows
histograms of the Spearman correlation (Spearman's $\rho$, a rank-based measure
of association) between the $\WIS(F)/\WIS(AR)$ and the magnitude of the
swing. Again we see that case rate increases are positively related to
diminished performance of the indicator models.

>>>>

As explained in Section 2.B of the main paper, a (somewhat cynical) hypothesis
for why we see benefits in forecasting and hotspot prediction is that the
indicators are not actually providing useful information but they are instead
acting as a sort of ``implicit regularization,'' leading to shrinkage on the
autoregressive coefficients and therefore to less volatile predictions.  To
investigate this hypothesis, we consider fitting ``noise features'' that in
truth have zero relationship to the response.  Recall (from the main paper) that
at each forecast date, we train a model on 6,426 location-time pairs.  Each
indicator model uses 6 features, corresponding to the 3 autoregressive terms and
the 3 lagged indicator values.  To form noise indicator features, we replace
their values with those from a randomly chosen time-space pair (while keeping
the autoregressive features fixed).  In particular, at each location $\ell$ and
time $t$, for the forecasting task we replace the triplet \smash{$(X_{\ell,t},
X_{\ell,t-7}, X_{\ell,t-14})$} in Eq.\ (3) of the main
paper with the triplet \smash{$(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
  X_{\ell^*,t^*-14})$}, where $(\ell^*,t^*)$ is a location-time pair sampled with
replacement from the 6,426 total location-time pairs.  Likewise in the hotspot 
prediction task, we replace the triplet \smash{$(X_{\ell,t}^\Delta,
  X_{\ell,t-7}^\Delta, X_{\ell,t-14}^\Delta)$} in Eq.\ (5) of the main paper with
\smash{$(X_{\ell^*,t^*}^\Delta, X_{\ell^*,t^*-7}^\Delta,
  X_{\ell^*,t^*-14}^\Delta)$}.  Figures
\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No method
exhibits a noticeable performance gain over the \ar~method, leading us to
dismiss the implicit regularization hypothesis.

@@

In this section, we provide extra details about the up/down/flat analysis
described in the main manuscript. Figure \ref{fig:upswing-summary} shows the
overall results, examining the average difference $\WIS(AR) - \WIS(F)$ in
period. Figure \ref{fig:hotspots-upswing-downswing} shows the same information
for the hotspot task. On average, during downswings and flat periods, the
indicator models have lower classification error and higher log likelihood than
the AR model. For hotspots, both Google-AA and CTIS-CLIIC perform better than
the AR model during upswings, in contrast to the forecasting task, where only
Google-AA improves. For a related analysis, Figure \ref{fig:cor-wis-ratio} shows
histograms of the Spearman correlation (Spearman's $\rho$, a rank-based measure
of association) between the $\WIS(F)/\WIS(AR)$ and the magnitude of the
swing. Again we see that case rate increases are positively related to
diminished performance of the indicator models.

### 2021-10-31 forecast/paper/supplement.Rmd
Copyedits to discussion

```{r lagging-only, fig.cap="Correlation of the difference in WIS with the laggingness of the indicator at the target date, stratified by up, down, or flat period. Compare to Figure 5 in the manuscript."}
# Deprecated
df2 %>%
  group_by(forecaster, udf) %>%
  summarise(
    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, laggingness, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = fcast_colors2, guide = guide_legend(nrow = 1)) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

@@

```{r diff-in-lead-lag, fig.cap = "Correlation of the difference between leadingness and laggingness with the difference in WIS. The relationship is essentially the same as described in the manuscript and shown in Figure 5."}
# Deprecated
df2 %>%
  group_by(forecaster, udf) %>%
  summarise(cor = cor(AR_wis - wis, leading - lagging, 
                      use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, cor, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = fcast_colors2) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

>>>>

```{r lagging-only, fig.cap="Correlation of the difference in WIS with the laggingness of the indicator at the target date, stratified by up, down, or flat period. Compare to Figure 5 in the manuscript."}
df2 %>%
  group_by(forecaster, udf) %>%
  summarise(
    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, laggingness, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = fcast_colors2, guide = guide_legend(nrow = 1)) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

@@

```{r diff-in-lead-lag, fig.cap = "Correlation of the difference between leadingness and laggingness with the difference in WIS. The relationship is essentially the same as described in the manuscript and shown in Figure 5."}
df2 %>%
  group_by(forecaster, udf) %>%
  summarise(cor = cor(AR_wis - wis, leading - lagging, 
                      use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, cor, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = fcast_colors2) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```
